Carbonate chemistry at Ny-Ålesund

### 2021-11-25 awipev-CO2.Rmd
We changed the selecting parameter "temp_insitu_ctd" from to "temp_mix_interp" to create pHseafet df from "d". we created before pHinsi() an interpolation for sal_mix/temp_mix in order to fill gaps for pHinsi(). For pHinsi(), we changed the sal/temp to sal_mix_interp/temp_mix_interp. We forced the depth in the function to 11 m.

---
title: "Carbonate chemistry at Ny-Ålesund" 
author: "Jean-Pierre Gattuso and Samir Alliouane"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  html_document: default
  word_document: default
  pdf_document: default
  odt_document: default
---
![ ](Images/AWIPEV-Logo.jpg){width=250px}

@@

```{r read observatory data, cleanup and prepare ferrybox data, echo=FALSE, warning=FALSE, message=FALSE}
d_hour <-
  readRDS(file = paste0(
    path,
    "fb_awipev-co2_server/ny-alesund/data/NRT_data/nydata_hour.rds"
  ))

@@

d_all <- as_tsibble(d_hour, index = datetime) #d_all will be the merger
d_all <- d_all %>%
  dplyr::mutate_all(function(x) ifelse(is.nan(x), NA, x)) %>% # replace all NaN by NA
  dplyr::mutate(depth_group_pressure = ifelse( # create new variable depth
      pressure_insitu_ctd <= 2,
      1,
      ifelse(
        pressure_insitu_ctd > 2 & depth <= 4,
        3,
        ifelse(
          pressure_insitu_ctd > 4 & depth <= 6,
          5,
          ifelse(
            pressure_insitu_ctd > 6 & depth <= 8,
            7,
            ifelse(pressure_insitu_ctd > 8, 10, NA)
          )
        )
      )
    ),
    depth_group_philipp = ifelse( # create new variable depth
      depth <= 2,
      1,
      ifelse(
        depth > 2 & depth <= 4,
        3,
        ifelse(
          depth > 4 & depth <= 6,
          5,
          ifelse(
            depth > 6 & depth <= 8,
            7,
            ifelse(depth > 8, 10, NA)
          )
        )
      )
    )) %>%
  dplyr::rowwise() %>% 
  dplyr::mutate(n_depth_ctd = sum(is.na( # create new variable to identify the hours for which there are CTD data at more than 1 depth 
  c(
    sal_insitu_ctd_1m,
    sal_insitu_ctd_3m,
    sal_insitu_ctd_5m,
    sal_insitu_ctd_7m,
    sal_insitu_ctd_9m
  )
), na.rm = TRUE)) %>% 
  ungroup()
#restore datetime, date which are mangled by the mutates above, arrange ascending
d_all <- d_all %>% 
  dplyr::mutate(datetime = d_hour$datetime, 
                date = d_hour$date,
                year = lubridate::year(d_hour$datetime), # year as a number
                monthd = lubridate::month(d_hour$datetime), # month as a number
                montht = lubridate::month(d_hour$datetime, label = TRUE, abbr = TRUE) # abbreviated months
  )
rm(d_hour)
```

@@

# add sal_fb + sal_insitu_ctd + temp_fb + temp_insitu_ctd from d_all into d, to use pHinsi
# use the closest time to add them
closest <- which.closest(d$datetime, d_all$datetime) # closest date and time
d$closest_datetime <- d_all$datetime[closest]
# if the closest value is more than 30 min away, then we do not take it
d$closest_datetime[abs(d$datetime - d_all$datetime[closest]) > 30*60] <- NA
# create closest_datetime in d_all (copy of datetime)
d_all$closest_datetime <- d_all$datetime
d <- left_join(d, d_all %>%
                 dplyr::select(
                   c(closest_datetime,
                     sal_fb,
                     temp_fb,
                     temp_mix,
                     sal_mix,
                     temp_insitu_11m,
                     temp_insitu_ctd,
                     depth
                   )
                 ), by = "closest_datetime") %>% 
  dplyr::select(-closest_datetime)

# Conversion of spec pH of in situ samples to in situ temp and 4 °C
pH_seafet <- d %>%
  dplyr::filter(!is.na(ph_s_sf) & !is.na(temp_insitu_ctd))
pH_seafet <- pH_seafet %>% 
  dplyr::mutate(
    ph_s_sf_temp_insi = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = pH_seafet$temp_insitu_ctd, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$depth/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),
   ph_s_sf_temp_4 = pHinsi(pH = pH_seafet$ph_s_sf, # with Lueker
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$depth/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
  )
# Conversion of spec pH of FerryBox samples to FerryBox temp
pH_durafet <- d %>%
  tidyr::drop_na(ph_s_dur, sal_fb, temp_fb)
pH_durafet <- pH_durafet %>% 
  dplyr::mutate(
    ph_s_dur_temp_fb = pHinsi(pH = pH_durafet$ph_s_dur, 
                              ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = pH_durafet$temp_fb, 
                              Tlab = pH_durafet$temp_s_lab, Pinsi = 0, S = pH_durafet$sal_fb, 
                              Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74"),
   ph_s_dur_temp_4 = pHinsi(pH = pH_durafet$ph_s_dur, # lueker
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4,
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0,
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d",
                         pHscale = "T", b="u74")#,
   # ph_s_dur_temp_4_l = pHinsi(pH = pH_durafet$ph_s_dur, # lueker
   #                       ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4,
   #                       Tlab = pH_durafet$temp_s_lab, Pinsi = 0,
   #                       S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d",
   #                       pHscale = "T", b="u74"),
   # ph_s_dur_temp_4_s20 = pHinsi(pH = pH_durafet$ph_s_dur, # Sulpis
   #                       ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4,
   #                       Tlab = pH_durafet$temp_s_lab, Pinsi = 0,
   #                       S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "s20", kf = "dg", ks="d",
   #                       pHscale = "T", b="u74"),       # Millero 02
   # ph_s_dur_temp_4_p18 = pHinsi(pH = pH_durafet$ph_s_dur, # Papadimitriou
   #                       ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4,
   #                       Tlab = pH_durafet$temp_s_lab, Pinsi = 0,
   #                       S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "p18", kf = "dg", ks="d",
   #                       pHscale = "T", b="u74"),
   # ph_s_dur_temp_4_m02 = pHinsi(pH = pH_durafet$ph_s_dur,  # Millero 02
   #                       ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4,
   #                       Tlab = pH_durafet$temp_s_lab, Pinsi = 0,
   #                       S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "m02", kf = "dg", ks="d",
   #                       pHscale = "T", b="u74")

>>>>

---
title: "Carbonate chemistry at Ny-Ålesund" 
author: "Jean-Pierre Gattuso and Samir Alliouane"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  html_document: default
  word_document: default
  pdf_document: default
  odt_document: default
editor_options: 
  chunk_output_type: console
---
![ ](Images/AWIPEV-Logo.jpg){width=250px}

@@

```{r read observatory data, cleanup and prepare ferrybox data, echo=FALSE, warning=FALSE, message=FALSE}
d_hour <-
  readRDS(file = paste0(
    path,
    "fb_awipev-co2_server/ny-alesund/data/NRT_data/nydata_hour.rds"
  ))
d_hour <- as_tsibble(d_hour, index = datetime) %>%#d_all will be the merger
  tsibble::fill_gaps(.full=TRUE)
d_hour$date <-  as.Date(d_hour$datetime, format="%Y-%m-%d")

@@

d_all <- d_hour %>%
  dplyr::mutate_all(function(x) ifelse(is.nan(x), NA, x)) %>% # replace all NaN by NA
  dplyr::mutate(depth_group_pressure = ifelse( # create new variable depth
      pressure_insitu_ctd <= 2,
      1,
      ifelse(
        pressure_insitu_ctd > 2 & depth <= 4,
        3,
        ifelse(
          pressure_insitu_ctd > 4 & depth <= 6,
          5,
          ifelse(
            pressure_insitu_ctd > 6 & depth <= 8,
            7,
            ifelse(pressure_insitu_ctd > 8, 10, NA)
          )
        )
      )
    ),
    depth_group_philipp = ifelse( # create new variable depth
      depth <= 2,
      1,
      ifelse(
        depth > 2 & depth <= 4,
        3,
        ifelse(
          depth > 4 & depth <= 6,
          5,
          ifelse(
            depth > 6 & depth <= 8,
            7,
            ifelse(depth > 8, 10, NA)
          )
        )
      )
    )) %>%
  dplyr::rowwise() %>% 
  dplyr::mutate(n_depth_ctd = sum(is.na( # create new variable to identify the hours for which there are CTD data at more than 1 depth 
  c(
    temp_insitu_ctd_1m,
    temp_insitu_ctd_3m,
    temp_insitu_ctd_5m,
    temp_insitu_ctd_7m,
    temp_insitu_ctd_9m
  )
), na.rm = TRUE)) %>% 
  ungroup()
#restore datetime, date which are mangled by the mutates above, arrange ascending
d_all <- d_all %>% 
  dplyr::mutate(datetime = d_hour$datetime, 
                date = d_hour$date,
                year = lubridate::year(d_hour$datetime), # year as a number
                monthd = lubridate::month(d_hour$datetime), # month as a number
                montht = lubridate::month(d_hour$datetime, label = TRUE, abbr = TRUE) # abbreviated months
  )
rm(d_hour)

@@

# interpolation of temp_mix/sal_mix in d_all in order to have a full temperature for pHinsi()
# We take temp_mix/sal_mix because of the depth (temp11m insitu + temp9m ctd)
tempmixinterp <-approx(d_all$datetime, d_all$temp_mix, xout=d_all$datetime, method="linear", rule=2)
names(tempmixinterp) <- c("datetime", "temp_mix_interp")
tempmixinterp <- as.data.frame(tempmixinterp)
d_all$temp_mix_interp <- tempmixinterp$temp_mix_interp

salmixinterp <-approx(d_all$datetime, d_all$sal_mix, xout=d_all$datetime, method="linear", rule=2)
names(salmixinterp) <- c("datetime", "sal_mix_interp")
salmixinterp <- as.data.frame(salmixinterp)
d_all$sal_mix_interp <- salmixinterp$sal_mix_interp

# ta_xts <- dplyr::select(d_all, datetime, sal_mix, sal_mix_interp)
# ta_xts <- as.xts(ta_xts, order.by = ta_xts$datetime)
# dygraph(ta_xts, group = "awipev", main="sal mix", ylab="temp") %>%
#   dySeries("sal_mix", color = "black", pointSize =4, strokeWidth = 1, label = "ta") %>%
#   dySeries("sal_mix_interp", color = "blue", strokeWidth = 1, label = "ta interp") %>%
#   dyHighlight(highlightCircleSize = 8,highlightSeriesBackgroundAlpha =0.2,hideOnMouseOut =TRUE) %>%
#       dyOptions( drawGrid = TRUE, drawPoints = TRUE,useDataTimezone = TRUE) %>%
#     dyAxis("y",valueRange = c(33, 36)) %>%
#     dyLegend(show = "follow")   %>%
#       dyRangeSelector(height = 30, dateWindow= NULL)

# add sal_fb + sal_insitu_ctd + temp_fb + temp_insitu_ctd from d_all into d, to use pHinsi
# use the closest time to add them
closest <- which.closest(d$datetime, d_all$datetime) # closest date and time
d$closest_datetime <- d_all$datetime[closest]
# if the closest value is more than 30 min away, then we do not take it
d$closest_datetime[abs(d$datetime - d_all$datetime[closest]) > 30*60] <- NA
# create closest_datetime in d_all (copy of datetime)
d_all$closest_datetime <- d_all$datetime
d <- left_join(d, d_all %>%
                 dplyr::select(
                   c(closest_datetime,
                     sal_fb,
                     temp_fb,
                     temp_mix,
                     temp_mix_interp,
                     sal_mix,
                     sal_mix_interp,
                     temp_insitu_11m,
                     temp_insitu_ctd,
                     temp_sf,
                     depth
                   )
                 ), by = "closest_datetime") %>% 
  dplyr::select(-closest_datetime)

# Conversion of spec pH of in situ samples to in situ temp and 4 °C
pH_seafet <- d %>%
  dplyr::filter(!is.na(ph_s_sf) & !is.na(temp_mix_interp))

pH_seafet <- pH_seafet %>% 
  dplyr::mutate(
    ph_s_sf_temp_insi = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = pH_seafet$temp_mix_interp, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = 11/10, 
                         S = pH_seafet$sal_mix_interp, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),
   ph_s_sf_temp_4 = pHinsi(pH = pH_seafet$ph_s_sf, # with Lueker
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = 11/10, 
                         S = pH_seafet$sal_mix_interp, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
  )
# Conversion of spec pH of FerryBox samples to FerryBox temp
pH_durafet <- d %>%
  tidyr::drop_na(ph_s_dur, sal_fb, temp_fb)
pH_durafet <- pH_durafet %>% 
  dplyr::mutate(
    ph_s_dur_temp_fb = pHinsi(pH = pH_durafet$ph_s_dur, 
                              ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = pH_durafet$temp_fb, 
                              Tlab = pH_durafet$temp_s_lab, Pinsi = 0, S = pH_durafet$sal_fb, 
                              Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74"),
   ph_s_dur_temp_4 = pHinsi(pH = pH_durafet$ph_s_dur, # lueker
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4,
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0,
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d",
                         pHscale = "T", b="u74")#,
   # ph_s_dur_temp_4_l = pHinsi(pH = pH_durafet$ph_s_dur, # lueker
   #                       ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4,
   #                       Tlab = pH_durafet$temp_s_lab, Pinsi = 0,
   #                       S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d",
   #                       pHscale = "T", b="u74"),
   # ph_s_dur_temp_4_s20 = pHinsi(pH = pH_durafet$ph_s_dur, # Sulpis
   #                       ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4,
   #                       Tlab = pH_durafet$temp_s_lab, Pinsi = 0,
   #                       S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "s20", kf = "dg", ks="d",
   #                       pHscale = "T", b="u74"),       # Millero 02
   # ph_s_dur_temp_4_p18 = pHinsi(pH = pH_durafet$ph_s_dur, # Papadimitriou
   #                       ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4,
   #                       Tlab = pH_durafet$temp_s_lab, Pinsi = 0,
   #                       S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "p18", kf = "dg", ks="d",
   #                       pHscale = "T", b="u74"),
   # ph_s_dur_temp_4_m02 = pHinsi(pH = pH_durafet$ph_s_dur,  # Millero 02
   #                       ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4,
   #                       Tlab = pH_durafet$temp_s_lab, Pinsi = 0,
   #                       S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "m02", kf = "dg", ks="d",
   #                       pHscale = "T", b="u74")
